---
title: "R Notebook"
output:
  pdf_document: default
  html_notebook: default
---
# Problem 1
It is known that the following Leibniz series converages to $\frac{\pi}{4}$ as 
$n\to\infty$:
$$
S(n) = \sum_{k=0}^{n} (-1) ^ k \frac{1}{2k+1}
$$
Plot the difference between $\frac{\pi}{4}$ sum $S(n)$ versus $n$ for
$0 \leq n \leq 200$.

```{r}
calculate_S <- function(n) {
  total <- 0
  for(k in 0:n){
    total <- total + (-1) ** k * (1 / (2*k+1))
  }
  return(total)
}

n <- 0:200
S = sapply(n, calculate_S)
plot(n, S - pi/4)
``` 

# Problem 2
The Virginia Cooperative Extension reports that the mean weight of yearling
Angus steers is 1,152 pounds. Suppose that weights of all such animals can be
described by a Normal model with a standard deviation of 84 pounds.

Plot the normal distribution curve ($\mu = 1152$ and $\sigma = 84$).  Fill the
distribution curve with your choice of color where:
  a. Weight is over 1250 pounds
  b. Weight is under 1200 pounds
  c. Weight is between 1000 and 1100 pounds
  
```{r}
x <- seq(900, 1400, by = 10)
data <- dnorm(x, mean=1152, sd=84)
plot(x, data, type="l")

polygon(
  c(x[x>=1250], 1250),
  c(data[x>=1250], min(data)),
  col="red"
)
polygon(
  c(x[x<=1200], 1200),
  c(data[x<=1200], min(data)),
  col="blue"
)
polygon(
  c(1000, x[x>=1000 & x<=1100], 1100),
  c(min(data), data[x>=1000 & x<=1100], min(data)),
  col="green"
)
```

# Problem 3
Read the "temperature.txt" file.  This file contains the following data:
  * High and low temperature (degree centigrade)
  * Rainfall (cm)
  * Month (from 1 to 12)
  * Year (for 20 years - from 1987 - 2005)
  
Create a boxplot of 'high temperature' using 'month' as a factor variable.
```{r}
temps <- read.table("temperatures.txt", header=TRUE)
month <- factor(temps$month)
plot(month, temps$temperature)
```

# Problem 4
In a binomial experiment with $n$ trials, the probability of success is $p$ in
each trial.  The probability of exactly $x$ success in $n$ trials is given
by the following expression:
$$
P(x:successes) = (_nC_xp^x(1-p)^{(n-x)})
$$
Suppose the probability of success $(p) = 0.7$ (probability of failure is
$(1 - 0.7) = 0.3$).  Therefore, in 8 trials ($n = 8$), the probability of
exactly 5 successes ($x=5$) is:
$$
=_8C_5(0.7)^5(0.3)^{(6-5)} = 0.25412
$$
Vary the probability of success from 0.1 to 0.9 and generate the following 9
barplots in R that represent the probability of$x$ successes in $8$ trials.
```{r}
prob_succ <- function(num_succ, prob) {
  return(
    choose(8, num_succ) * prob ** num_succ * 0.3 ** (8-num_succ)
  )
}
par(mfrow=c(3,3))
for (p in 1:9){
  c = 1:8
  barplot(
    sapply(c, prob_succ, prob=p/10), 
    xlab="Trials",
    ylab="Probability"
    )
}

```
