---
title: "Homework 6"
author: "Eric Neer"
output:
  pdf_document: 
    fig_caption: yes
    keep_tex: yes
  html_notebook: default
header-includes: \usepackage{float}
---

# Linear Regression
The values of 2 variables X and Y are given below. Here X is the predictor
variable and Y is the response variable.

```{r}
X <- c(25, 30, 35, 40, 45)
Y <- c(5, 260, 480, 745, 1100)
```
## Problem 1
Compute the correlation between X and Y variables. Verify your answer using
Excel.
```{r}
correlation = cor(X,Y)
```

## Problem 2
Compute regression equation using Excel by using a scatter plot. Compute the
value of RSquare.  

\begin{figure}[h!]
  \caption{Excel Chart Generated Using X, Y Data}
  \centering
    \includegraphics[]{../Homework_6/p2.png}
\end{figure}
  
## Problem 3
Compute regression equation using the following parameters:
* Correlation between X and Y variables
* Mean and the standard deviation of X and Y variables
```{r}
slope = correlation * sd(Y) / sd(X)
intercept = mean(Y) - slope * mean(X)
```
Thus, the regression equation is (using $y = mx + b$):
$$
y = 53.5x - 1354.5
$$

## Problem 4
Compute regression equation using Excel/Solver by minimizing the square of the
difference. Compute the values of J, S, and RSquare.  

\begin{figure}[h!]
  \caption{Parameters Calculated with Excel}
  \centering
    \includegraphics[]{../Homework_6/p4.png}
\end{figure}


## Problem 5
Compute regression equation using R software.
```{r}
result <- lm(Y~X)
```

## Problem 6

Prediction: Predict the value of Y variable when the value of X = 40 and 56
using the regression equation computed.
```{r}
new.df <- data.frame(X=c(40, 56))
predict(result, new.df)
```
# Multiple Regression

## Problem 7
Compute regression equation using Excel/Solver by minimizing the square of the
difference. Compute the value of RSquare.

\begin{figure}[h!]
  \caption{Parameters Calculated with Excel}
  \centering
    \includegraphics[]{../Homework_6/p7.png}
\end{figure}

## Problem 8
Compute regression equation using R software. Compute the value of RSquare.
```{r}
p8data <- read.csv("../Homework_6/P7-9RawData.csv", header=TRUE)
attach(p8data)
result <- lm(
  Revenue ~ NewVisits + PagesPerSession + AvgSessionDuration
  )
summary(result)$r.squared
```

## Problem 9
Predict the revenues for a period where:
  * New Vists = 32
  * Page per Session = 7
  * Average Session Duration = 457
```{r}
new.df <- data.frame(NewVisits=32, PagesPerSession=7, AvgSessionDuration=457)
predict(result, new.df)
```