---
title: "Homework 2"
author: "Eric Neer"
date: "October 1, 2017"
output:
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Problem 1
Given x and y
```{r}
x <- 5 + 8i
y <- -6 + 7i
```
Evaluate the following expressions in R.
```{r}
  (u <- x + y)
  (v <- x * y)
  (w <- x / y)
  (z <- exp(x))
  (r <- sqrt(y))
  (s <- x * y ** 2)
```
## Problem 2
Use R to compute the following expressions.
```{r}
  (a2 <- (3 + 6i)*(-7 - 9i))
  (b2 <- (5 + 4i) / (5 - 4i))
  (c2 <- 3 / 2i)
```
## Problem 3
Use R to calculate the following expressions.
```{r}
  (a3 <- exp(-2.1) ** 3 + 3.47 * log(14) + 287 ** (1/4))
  (b3 <- 3.4 ** 7 * log(14) + 287 ** (1/4))
  (c3 <- cos((4.12 * pi) / 6) ** 2)
  (d3 <- cos((4.12 * pi) / 6) ** 2)
```
## Problem 4
Suppose that x = 6. Find the results of the following operations by hand 
and use R to check your results.
```{r}
  x <- 6
  (a4 <- (x < 10))
  (b4 <- (x == 10))
  (c4 <- (x >= 4))
  (d4 <- (x != 7))
```
## Problem 5
Find the results of the following operations by hand and use R to check
your results.

```{r}
  (a5 <- 6 > 3 + 8)
  (b5 <- 6 + 3 > 8)
  (c5 <- 4 > (2 + 9))
  (d5 <- (4 < 7) + 3)
  (e5 <- 4 < 7 + 3)
  (f5 <- (4 < 7) * 5)
  (g5 <- 4 < (7 * 5))
  (h5 <- 2 / 5 >= 5)
```
## Problem 6
Enter this matrix in R:
$$
A = \left[\begin{array}
{ccc}
3 & 7 & -4 & 12 \\
-5 & 9 & 10 & 2 \\
6 & 13 & 8 & 11 \\
15 & 5 & 4 & 1
\end{array}\right]
$$

```{r}
  (A <- matrix(
        c(
          3, 7, -4, 12,
          -5, 9, 10, 2,
          6, 13, 8, 11,
          15, 5, 4, 1
          ),
        nrow=4,
        byrow=TRUE
  ))
```
a) Create a vector V consisting of the elements in the second column of A.
```{r}
(v <- A[,2])
```
b) Create a vector W consisting of the elements in the second row of A.
```{r}
(w <- A[2,])
```
## Problem 7
Store 20 random numbers with normal distribution and mean of 10 and
standard deviation of 5 in a vector. Sort the vector. Remove the smallest 2
and the largest 2 elements from this vector.
```{r}
  (B <- rnorm(20, mean=10, sd=5))
  
  trim <- function(x) {
    last_idx = length(x)
    return(sort(x)[-c(1, 2, last_idx-1, last_idx)])
  }
  
  (trim(B))
```
## Problem 8
Read the 'temperature.txt' file. This file contains the following data.
   - High and low temperature (degree centigrade)
   - Rainfall (centimeters)
   - Month (from 1 to 12)
   - Year (for 20 years - from 1987 - 2005)

 Compute the following statistics of temperature and rainfall data.
   Mean, Max, Min, Standard deviation, Variance, Range
```{r}
  temps <- read.table("temperatures.txt", header=TRUE)
  funcs <- list(mean=mean, max=max, min=min, sd=sd, var=var)
  (temp_stats <- sapply(funcs, mapply, list(temps$temperature, temps$rain)))
  (range_temp <- range(temps$temperature))
  (range_rain <- range(temps$rain))
```
## Problem 9
Using the text file of problem 8, compute the yearly average temperature
for all the years (from 1987 - 2005). Which was the coldest year (year in
which the average temperature was the lowest) and which was the warmest?
```{r}
  (yearly_avg <- tapply(temps$temperature, temps$yr, mean))
  (coldest_yr <- names(sort(yearly_avg)[1]))
  (warmest_yr <- names(sort(yearly_avg)[length(yearly_avg)]))
```
## Problem 10
Store 100 random numbers with normal distribution and mean of 50 and
standard deviation of 50 in a vector. Plot this histogram. (you will see the
bell shaped Gaussian curve). Which number is the closest to the number 100?
```{r}
  C <- rnorm(100, mean=50, sd=50)
  hist(C)
  
  # get the index of the closest number and the closest number itself
  (closest_idx <- which (abs(C-100) == min(abs(C-100))))
  (closest_num <- C[closest_idx])
```