---
title: "R Notebook"
output:
  pdf_document: default
  html_notebook: default
---

# Problem 1
The half-life of a radioactive substance is the time it takes to decay by
one-half.  The half-life of carbon 14, which is used for dating previously
living things, is 5500 years.  When an organism dies, it stops accumulating
carbon 14.  The carbon 14 present at the time of death decays with time.  Let
$\frac{C(t)}{C(0)}$ be the fraction of carbon 14 remaining at time $t$.  In
radioactive carbon dating, scientists usually assume that the remaining fraciton
decays exponentially according to the following formula:
$$
\frac{C(t)}{C(0)} = e^{-bt}
$$

  a.  Use the half-life of carbon 14 to find the value of the parameter $b$ and
  plot the function.
  
```{r}
# at time t = 5500, half of the carbon remains.  i.e. C(5500)/C(0) = 0.5
b <- log(0.5) / -5500
t <- seq(0, 5e4, 100)
carbon_fraction <- function(t) {
  return(exp(-b*t))
}
plot(
  t, 
  carbon_fraction(t), 
  type="l", 
  ylab="Carbon Fraction Remaining",
  xlab="Time Elapsed (years)"
  )
```
  
  b.  If 90% of the original carbon 14 remains, estimate how long ago the 
  organism died.

```{r}
# C(t) = 0.9, b found in part a
years_since_death <- log(0.9) / -b
```

# Problem 2
Given the matrices:

$$
A = \left[\begin{array}
{ccc}
4 & -2 & 1 \\
6 & 8 & -5 \\
7 & 9 & 10
\end{array}\right]
$$
$$
B = \left[\begin{array}
{ccc}
6 & 9 & -4 \\
7 & 5 & 3 \\
-8 & 2 & 1 
\end{array}\right]
$$
$$
C = \left[\begin{array}
{ccc}
-4 & -5 & 2 \\
10 & 6 & 1 \\
3 & -9 & 8
\end{array}\right]
$$
use R to verify:

  a.  The associative property $\bf{A(B+C) = AB+AC}$
```{r}
A <- matrix(c(4, -2, 1, 6, 8, -5, 7, 9, 10), byrow=TRUE, nrow=3)
B <- matrix(c(6, 9, -4, 7, 5, 3, -8, 2, 1), byrow=TRUE, nrow=3)
C <- matrix(c(-4, -5, 2, 10, 6, 1, 3, -9, 8), byrow=TRUE, nrow=3)

identical(A %*% (B + C), A %*% B + A %*% C)
```

  b. The distributive property
```{r}
identical((A %*% B) %*% C, A %*% (B %*% C))
```

# Problem 3
Suppose it is known that the graph of the following function
$$
y = ax^3 + bx^2 + cx + d
$$
passes through four given points $(x_i, y_i)$ where $i = 1, 2, 3, 4$.

Write a user defuned function that accepts these four points as input and
computes the coefficients $a, b, c, d$.  The function should solve four linear
equations in terms of the four unknown coefficients.  Test your function for the
case where $(x_i, y_i) = $ (-2,-20), (0, 4), (2, 68) and (4, 508).

The answer is $a = 7$, $b = 5$, $c = -6$, $d = 4$.
```{r}
calculate_coefficients <- function(point1, point2, point3, point4) {
  A <- matrix(c(
    point1[1]**3, point1[1]**2, point1[1], 1,
    point2[1]**3, point2[1]**2, point2[1], 1,
    point3[1]**3, point3[1]**2, point3[1], 1,
    point4[1]**3, point4[1]**2, point4[1], 1
  ), nrow=4, byrow=TRUE)

  b <- matrix(c(
    point1[2], point2[2], point3[2], point4[2]),
    nrow=4
  )
  return(solve(A, b))
}
p1 <- c(-2, -20)
p2 <- c(0, 4)
p3 <- c(2, 68)
p4 <- c(4, 508)
calculate_coefficients(p1, p2, p3, p4)
```

# Problem 4
Use a random number generator to produce 1000 uniformly distributed numbers with
a mean of 10, a minimum of 2, and a maximum of 18.  Obtain the mean and the
histogram of these numbers, and discuss whether they appear uniformly
distributed with the desired mean.

```{r}
r_uni <- runif(1000, min=2, max=18)
mean(r_uni)
```
```{r}
hist(r_uni)
```

## Discussion
The dataset is not uniformly distributed nor is the mean 10 as desired.  This
is due to the fact that this is only a random sample of 1000 numbers
($n_{sample} = 1000$) from a uniform distribution.  As more data points are
taken, the histogram will show a more evenly spread distribution and the mean
will approach 10.  In the limiting case of $n_{sample} \to \infty$, the
histogram would show a perform uniform distribution and the mean would be
exactly 10.

# Problem 5

Use a random number generator to produce 1,000 normally distributed numbers with
a mean of 20 and a standard deviation of 4. Obtain the mean, standard deviation,
and the histogram of these numbers, and discuss whether they appear normally
distributed with the desired mean and standard deviation.

```{r}
r_norm <- rnorm(1000, mean=20, sd=4)
mean(r_norm)
```
```{r}
hist(r_norm)
```

## Discussion

Similar to the uniform distribution, only a 1000 point sample is shown, thus
the histogram and mean do not reflect a perfect normal distribution.

# Problem 6
The mean of the sum (or difference) of two independent random variables equals
the sum (or difference) of their means, but the variance is always the sum of
the two variances.  

Use random number generation to verify this statement for the case where
$z = x + y$, where $x$ and $y$ are independent and normally distributed random
variables.  The mean and the variance of $x$ and $y$ are given below.
$$
\mu_x = 8 \quad \sigma_x^2 = 2 \\ 
\mu_y = 15 \quad \sigma_y^2 = 4
$$
```{r}
x <- rnorm(1e4, mean=8, sd=sqrt(2))
y <- rnorm(1e4, mean=15, sd=sqrt(4))
mean(x+y) == mean(x) + mean(y)
var(x+y) == var(x) + var(y)
```