---
title: "R Programming - Homework 3"
author: "Eric Neer"
date: "October 8, 2017"
output:
  pdf_document: default
  html_notebook: default
---
# Problem 1
Read a file 'blowfly.txt: Compute the total number of records in this file.
How many of them are unique?
```{r}
bf <- read.table('blowfly.txt', header=TRUE)
(total_records <- length(bf$flies))
(unique_records <- length(unique(bf$flies)))
```
# Problem 2
Use a loop in R to determine how long it will take to accumulate $1,000,000 in a
bank account if you deposit $10,000 initially and $10,000 at the end of each
year; the account pays 6% annual interest.
```{r}
balance <- 10e3
years <- 0

while (balance < 1e6){
  balance <- balance * 1.06 + 10e3
  years <- years+ 1
}

(years)
```
# Problem 3
Read the 'worldfloras.txt' file in R. Display the countries names that have a
character 'c' as the 4th character in their name. Use 'R 'grep' command and
regular expressions.
```{r}
wf <- read.table("worldfloras.txt", header=TRUE)
pattern <- "^.{3}c[a-zA-Z]*"
wf$Country[grep(pattern, wf$Country)]
```
#Problem 4
The volume $V$ and paper surface area $A$ of a conical paper cup are given by
the following 2 equations.
$$
V = \frac{1}{3}\pi r^2h
$$
$$
A = \pi r^2 \sqrt{r^2 + h^2}
$$
Where $r$ is the radius of the base of the cone and $h$ is the height of the cone.
By eliminating $h$, obtain the expression for $A$ as a function of $r$ and $V$.

Rearranging the equation for $V$:
$$
h = \frac{3V}{\pi r^2}
$$
And substituting it in to the equation for $A$:
$$
A = \pi r^2 \sqrt{r^2 + \left(\frac{3V}{\pi r^2}\right)^2} = \pi r^2 \sqrt{r^2 + \left(\frac{9V^2}{\pi^2 r^4}\right)}
$$
Create a user-defined function that accepts $r$ and $V$ as arguments and computes
$A$ for a given value of $V$ (assume $V = 10 in^3$).
```{r}
conical_area <- function(r, V=10){
  A = pi * r**2 * sqrt(r**2 + (9 * V**2) / (pi**2 * r**4))
  return(A)
}
```
Plot a graph between $A$ and $r$ where $r$ varies from 0.1 to 10 inches.
```{r}
radius <- seq(0.1, 10, 0.1)
plot(radius, 
     conical_area(radius),
     main="Conical Surface Area vs Radius",
     ylab="Surface Area",
     xlab="Radius"
     )
```
For which value of $r$, $A$ value is minimum?
```{r}
areas <- conical_area(radius)
min_idx = match(min(areas), areas)
(radius[min_idx])
```
#Problem 5
The recursive definition of Factorial function is as follows.
$$
Factorial(x) = x * Factorial(x-1)
$$
Write a recursive function R that computes the factorial of a number. Test that
function for all the numbers from 1 to 10.
```{r}
my_factorial <- function(x){
  if (x %in% c(0,1)){
    return(1)
  } else {
    return (
      x * my_factorial(x-1)
    )
  }
}

(sapply(seq(1,10), my_factorial))
```
#Problem 6
Read 'cells.txt' and 'multivariate.txt' files into R using 'read.table' command.
```{r}
cells <- read.table("cells.txt", header=TRUE)
multi <- read.table("multivariate.txt", header=TRUE)
```
## Part a
Using R functions identify the data type of each column for both files
```{r}
(sapply(cells, class))
(sapply(multi, class))
```
## Part b
Read both files again. This time omit the column headers and assign column names of your own choice.
```{r}
cells_names <- c("CELLS", "SMOKER", "AGE", "SEX", "WEIGHT")
multi_names <- c("TEMP", "INDUSTRUY", "POPULATION", "WIND", "RAIN", "WET_DAYS")

cells <- read.table("cells.txt", header=TRUE, col.names=cells_names)
multi <- read.table("multivariate.txt", header=TRUE, col.names=multi_names)

colnames(cells)
colnames(multi)
```
## Part c
Save the two datasets to both ASCII text using 'write.table' command 
```{r}
write.table(cells, "cells_data.txt")
write.table(multi, "multi_data.txt")
```
and binary dataframe files using 'save' command.
```{r}
cells_out <- file("cells_b", "wb")
multi_out <- file("multi_b", "wb")

save(cells, file=cells_out)
save(multi, file=multi_out)

close(cells_out)
close(multi_out)
```