---
title: "Homework 7"
author: "Eric Neer"
output:
  pdf_document: default
  html_notebook: default
  html_document: default
---
# Problem 1
Given data captured by Analytics service of a website, write R code to predict 
if the people with the following characteristics will visit again or not. 
Compute probabilities for each case. If the probability is greater than 0.5, 
visitor will visit again.

Load in the training and testing data, removing the primary key of each record.
```{r}
library(e1071)

train <- read.csv("HW07Train.csv")
train <- train[
  c("Previously_Visited", "SessionLength", "PageViews", "Will_Visit_Again")
  ]

test <- read.csv("HW07Test.csv")
test <- test[
  c("Previously_Visited", "SessionLength", "PageViews")
  ]
```

Use the naiveBayes class to generate a classifier for whether or not a visitor
will visit the given website again.
```{r}
classifier <- naiveBayes(train, train$Will_Visit_Again, laplace=0)
predictions <- predict(classifier, test, type="raw")
predictions
```
\begin{center}
  \begin{tabular}{c c c | c c | c}
    Previously Visited & Session Length & Page Views & No & Yes & Prediction \\
    \hline
    No & Short & Low & 0.624 & 0.376 & No \\ 
    Yes & Short & Medium & 0.740 & 0.260 & No \\ 
    Yes & Average & Medium & 0.850 & 0.150 & No \\ 
    Yes & Long & High & 0.453 & 0.547 & Yes \\ 
  \end{tabular}
\end{center}